A summary on Delta Debugging and its uses

### 2019-01-11 Paper/article.tex
Started on use case 2.

% Colors
\usepackage{xcolor}

@@

% Colored text shortcuts
\newcommand{\green}[1]{\textcolor{green}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\gray}[1]{\textcolor{lightgray}{#1}}

@@

@@

Consider as an example the C program \texttt{fail.c} which caused the GNU C Compiler (GCC) in version XY %TODO VERSION!
to crash. The program looks like correct C code that should be able to be compiled. We'd like to use Delta Debugging to find out which part of \texttt{fail.c} is responsible for the crash. For this an empty C program is regarded as the \yd state and \texttt{fail.c} as the \td state. A change is defined as a C language token. $\C$ is comprised of all tokens in \texttt{fail.c}. As depicted in \ref{fig:ddponfail} \ddp\ only takes 21 tests to find out that the C tokens "$+ 1.0$" are a minimal failure-inducing set of changes.
\\
\begin{figure}[h!]
	\fbox{
		\begin{minipage}{\textwidth}
			\begin{centering}
			\begin{tabular}{clc}
				\# & GCC input & test \\
				\hline
				1 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots\}$ & $\xmark$ \\
				2 &\gray{ double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots\}$} & $\cmark$ \\
				3 & double mult($\dots$) $\{\gray{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots}\}$ & $\cmark$ \\
				4 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; \gray{for (\dots) \{ \dots\} \dots}\}$ & $\cmark$ \\
				5 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\}\gray{ \dots}\}$ & $\xmark$ \\
				6 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\gray{\dots}) \{ \dots\} \gray{\dots}\}$ & $\cmark$ \\

				\vdots & \multicolumn{1}{c}{\vdots} & \vdots \\
				18 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] + 1.0);$ \quad \dots } & \xmark \\
				19 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] \gray{+ 1.0});$ \quad \dots } & \cmark \\
				20 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] \gray{+} 1.0);$ \quad \dots } & \qmark \\
				21 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] + \gray{1.0});$ \quad \dots } & \qmark \\
			\end{tabular}
			\end{centering}
		\end{minipage}
	}
	\label{fig:ddponfail}
	\caption{How \ddp\ finds the part of \texttt{fail.c} that is responsible for GCC crashing.}
\end{figure}

\subsection{Program runtime analysis}
\subsection{Minimizing random unit test cases}
\subsubsection{The ddmin algorithm}

>>>>

% Colors
\usepackage{xcolor}
\definecolor{td-green}{rgb}{0,.59,.51}

% Better references
\usepackage{hyperref}

@@

% Colored text shortcuts
\newcommand{\green}[1]{\textcolor{td-green}{#1}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\gray}[1]{\textcolor{lightgray}{#1}}

@@

% fail.c reference shortcut
\newcommand{\reffail}[0]{\href{fig:fail.c}{\ensuremath{fail.c^{Fig. \ref{fig:fail.c}}}}}

% r_pass and r_fail shortcuts
\newcommand{\rpass}[0]{\ensuremath{r_{\cmark}}}
\newcommand{\rfail}[0]{\ensuremath{r_{\xmark}}}

@@

Consider as an example the C program \reffail which caused the GNU C Compiler (GCC) in version 2.95.2 (on Intel-Linux with optimization enabled) to crash. The program looks like correct C code that should be compiled properly. We'd like to use Delta Debugging to find out which part of \reffail is responsible for the crash. An empty C program is regarded as the \yd state and \reffail as the \td state. A change is defined as a C language token. $\C$ is comprised of all tokens in \reffail. As depicted in \ref{fig:ddponfail} \ddp\ only takes 21 tests to find out that the C tokens "$+ 1.0$" are a minimal failure-inducing set of changes.
\\
\begin{figure}[h!]
	\fbox{
		\begin{minipage}{\textwidth}
		\begin{centering}
			\begin{algorithmic}[1]
					\State double \textbf{mult}(double z[], int n)
					\State $\{$
					\State \quad int i,j;
					\State \quad i = 0;
					\State \quad for(j = 0; j < n; j++) $\{$
					\State \quad \quad i = i + j + 1;
					\State \quad \quad z[i] = z[i] * (z[0] + 1.0);
					\State \quad $\}$
					\State \quad return z[n];
					\State $\}$
			\end{algorithmic}
		\end{centering}
		\end{minipage}
	}
	\caption{The $fail.c$ program that crashes GCC}
	\label{fig:fail.c}
\end{figure}
\\
\begin{figure}[h!]
	\fbox{
		\begin{minipage}{\textwidth}
			\begin{centering}
			\begin{tabular}{clc}
				\# & GCC input & test \\
				\hline
				1 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots\}$ & $\xmark$ \\
				2 &\gray{ double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots\}$} & $\cmark$ \\
				3 & double mult($\dots$) $\{\gray{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots}\}$ & $\cmark$ \\
				4 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; \gray{for (\dots) \{ \dots\} \dots}\}$ & $\cmark$ \\
				5 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\}\gray{ \dots}\}$ & $\xmark$ \\
				6 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\gray{\dots}) \{ \dots\} \gray{\dots}\}$ & $\cmark$ \\

				\vdots & \multicolumn{1}{c}{\vdots} & \vdots \\
				18 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] + 1.0);$ \quad \dots } & \xmark \\
				19 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] \gray{+ 1.0});$ \quad \dots } & \cmark \\
				20 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] \gray{+} 1.0);$ \quad \dots } & \qmark \\
				21 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] + \gray{1.0});$ \quad \dots } & \qmark \\
			\end{tabular}
			\end{centering}
		\end{minipage}
	}
	\caption{How \ddp\ finds the part of \texttt{fail.c} that is responsible for GCC crashing.}
	\label{fig:ddponfail}
\end{figure}

\subsection{Program runtime analysis}

It can be of great help to the software developer to be able to automatically determine the specific parts of the program input that cause the failure. The information gathered from the above method can be used by the developer to manually trace the responsible input data to the crash. Considering how a miniscule input change can cause entirely different execution paths this task is tedious if not impossible.
Delta Debugging can help trace the error from the root to the failure during program execution as well, though. Given a test framework that is able to access program states during execution, DD is used on the program states of two program runs \rpass\ and \rfail\ at significant locations to deduce a \textit{Cause-Effect-Chain (CEC)}.

\definition \defsub{Program states} A program state

The entire program execution is determined by a set of circumstances. We focus on changeable circumstances, more specifically the circumstances that influence the test outcome. A program run, i.e. one specific sequence of program states, is determined only by the set of circumstances that induce the specific run.

\definition \defsub{Circumstances and program runs} Let $R$ be the set of all possible circumstance configurations. A run is defined as a configuration of circumstances, so every $r \in R$ is called a run.

\subsection{Minimizing random unit test cases}
\subsubsection{The ddmin algorithm}
